---
title: "Patron_Lab2"
author: "Javier Patrón"
date: "2023-01-18"
output: pdf_document
---

Today we will be continuing the pumpkin case study from last week. We will be using the data that you cleaned and split last time (pumpkins_train) and will be comparing our results today to those you have already obtained, so open and run your Lab 1 .Rmd as a first step so those objects are available in your Environment (unless you created an R Project last time, in which case, kudos to you!).  

Once you have done that, we'll start today's lab by specifying a recipe for a polynomial model.  First we specify a recipe that identifies our variables and data, converts package to a numerical form, and then add a polynomial effect with step_poly()

```{r}
# Specify a recipe
poly_pumpkins_recipe <-
  recipe(price ~ package, data = pumpkins_train) %>%
  step_integer(all_predictors(), zero_based = TRUE) %>% 
  step_poly(all_predictors(), degree = 4)
```

How did that work? Choose another value for degree if you need to. Later we will learn about model tuning that will let us do things like find the optimal value for degree.  For now, we'd like to have a flexible model, so find the highest value for degree that is consistent with our data.

Polynomial regression is still linear regression, so our model specification looks similar to before.

```{r}
# Create a model specification called poly_spec
poly_spec <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")
```

Question 1: Now take the recipe and model specification that just created and bundle them into a workflow called poly_df.
```{r}
# Bundle recipe and model spec into a workflow
poly_wf <- workflow() |> 
  add_recipe(poly_pumpkins_recipe) |> 
  add_model(poly_spec)
  
```

Question 2: fit a model to the pumpkins_train data using your workflow and assign it to poly_wf_fit
```{r}

library(parsnip)

# Create a model
poly_wf_fit <- poly_wf |> 
  fit(data = pumpkins_train)
  
```

```{r}
# Print learned model coefficients
poly_wf_fit

```


```{r}
# Make price predictions on test data
poly_results <- poly_wf_fit %>% 
  predict(new_data = pumpkins_test) %>% 
  bind_cols(pumpkins_test %>% 
              select(c(package, price))) %>% 
  relocate(.pred, .after = last_col())

# Print the results
poly_results %>% 
  slice_head(n = 10)
```

Now let's evaluate how the model performed on the test_set using yardstick::metrics().
```{r}

metrics_linear <- metrics(data = lm_results, truth = price,estimate = .pred)
metrics_poly <- metrics(data = poly_results, truth = price, estimate = .pred)

print("Linear Regression Metrics:")
metrics_linear

print("Polynomial  Metrics:")
metrics_poly

```

Question 3: How do the performance metrics differ between the linear model from last week and the polynomial model we fit today?  Which model performs better on predicting the price of different packages of pumpkins?

*Response: As the result of the code chunk above, we can see that the polynomial model fits better than the linear regression here is why:*

*RMSE: Stands for "Root Mean Square Error", which tell you the aggregated magnitude of errors from the model for various data points, into a single measure of predictive power. RMSE is a measure of accuracy to compare forecasting errors of different models for a particular data set. A lower value of RMSE is better than a higher one. Comparing the polynomial (3.27) vs. the linear result (7.23) we can tell that the polynomial is less than half the value.*

*RSQ: Stands for Root Squared, where this coefficient of determination provides a measure for the predictions to the observations. This is a value between 0 and 1, where 0 means no-fit and 1 means perfect fit. In this case the polynomial result (0.892) is significantly higher, than the linear model (0.495).*

Let's visualize our model results.  First prep the results by binding the encoded package variable to them.

```{r}
# Bind encoded package column to the results
poly_results <- poly_results %>% 
  bind_cols(package_encode %>% 
              rename(package_integer = package)) %>% 
  relocate(package_integer, .after = package)


# Print new results data frame
poly_results %>% 
  slice_head(n = 5)
```

OK, now let's take a look! 

Question 4: Create a scatter plot that takes the poly_results and plots package vs. price.  Then draw a line showing our model's predicted values (.pred). Hint: you'll need separate geoms for the data points and the prediction line.

```{r}
# Make a scatter plot

my_formula <- price ~ package

poly_results %>% ggplot(aes(x=package_integer)) +
  geom_point(aes(y = price), color = "blue") +
  geom_point(aes(y=.pred), color = "red") +
  geom_line(aes(y=.pred), color = "gray30")

```

You can see that a curved line fits your data much better.

Question 5: Now make a smoother line by using geom_smooth instead of geom_line and passing it a polynomial formula like this:
geom_smooth(method = lm, formula = y ~ poly(x, degree = 3), color = "midnightblue", size = 1.2, se = FALSE)

```{r}
# Make a smoother scatter plot 

my_formula = y ~ poly(x,degree = 3)

poly_results %>% ggplot(aes(x= package_integer, y = price)) +
  geom_point(aes(y = price), color = "red") +
  geom_point(aes(y = .pred), color = "blue") +
  geom_smooth(method = lm,
              formula = my_formula,
              color = "gray25",
              se= F)
```

OK, now it's your turn to go through the process one more time.
 
Additional assignment components :
6. Choose  a new predictor variable (anything not involving package type) in this data set.
7. Determine its correlation with the outcome variable (price).  (Remember we calculated a correlation matrix last week)
8. Create and test a model for your new predictor:
  - Create a recipe
  - Build a model specification (linear or polynomial)
  - Bundle the recipe and model specification into a workflow
  - Create a model by fitting the workflow
  - Evaluate model performance on the test data
  - Create a visualization of model performance
  
  
```{r}



```

  
Lab 2 due 1/24 at 11:59 PM
